---
title: "HappinessAnalysis"
author: "Fatmah Kabeer"
date: "10/11/2020"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
library(tidyverse)
library(DT)
library(RColorBrewer)

```


In this document, I analyze a dataset that I have collected from Kaggle. This dataset illustrates the happiness in the world. I mainly focus on my analysis of the year 2019 as well as my country, Saudi Arabia.

```{r readingDate}
happiness_2015 <- read.csv("archive/2015.csv")
happiness_2016 <- read.csv("archive/2016.csv")
happiness_2017 <- read.csv("archive/2017.csv")
happiness_2018 <- read.csv("archive/2018.csv")
happiness_2019 <- read.csv("archive/2019.csv")
```


### In 2019


First of all we need to know the comparisons criterias in 2019:

```{r comparisonsCriteria}
dropN <- c("Country.or.region", "Overall.rank", "Score")
criteria <- names(happiness_2019[!(names(happiness_2019)) %in% dropN])

knitr::kable(criteria)
```


#### How many Countries in the evaluating list?
```{r}
nrow(happiness_2019)
```
<br />

#### What are the top 10 of the happiest countries in 2019?
```{r}
happiest <- arrange(happiness_2019, Overall.rank)
happiest10 <- head(happiest$Country.or.region, 10)

knitr::kable(happiest10)
```
<br />

#### The top 10 of the saddest countries in 2019:
```{r}
saddest <- arrange(happiness_2019, desc(Overall.rank))
saddest10 <- head(saddest$Country.or.region, 10) %>% 
   knitr::kable()
```


#### How freedom affect countries' happiness in 2019?

**Plot to explain the relationship between freedom to make life choices with the level of happiness.**
As we can see the population in the happiest countries is as showing kind of have a more degree of freedom to make life than others.

```{r freedomPlot}
ggplot(happiness_2019, aes(x = Country.or.region,
                           y = Overall.rank,
                           color = Freedom.to.make.life.choices)) +
  #facet_grid(Country.or.region ~ Social.support) +
  geom_point()
```
<br />
<br />

#### Comparison between the average Healthy Life Expectancy in the happiest against the saddest countries.
**The average of the Healthy Life Expectancy for the saddest country in 2019:**
```{r happinestAvg}
happiness_2019 %>% 
  summarise(avg = mean(Healthy.life.expectancy[Overall.rank > (nrow(happiness_2019) - 10)]))

```
**The average of the Healthy life expectancy for the happiest country in 2019:**
```{r saddnestAvg}
happiness_2019 %>% 
  summarise(avg = mean(Healthy.life.expectancy[Overall.rank > 10]))
```
<br />

### Saudi Arabia:
**Saudi Arabia rank**

```{r dataset for each year}
sa2015 <- filter(happiness_2015, Country == "Saudi Arabia")
sa2016 <- filter(happiness_2016, Country == "Saudi Arabia")
sa2017 <- filter(happiness_2017, Country == "Saudi Arabia")
sa2018 <- filter(happiness_2018, Country.or.region == "Saudi Arabia")
sa2019 <- filter(happiness_2019, Country.or.region == "Saudi Arabia")
```

```{r Saudi Arabia Rank in 2019}

saRank <- happiness_2019 %>% 
  filter(Country.or.region == "Saudi Arabia")
saRank[1,1]

```


#### Saudis Happiness details in 2019 in comparisons with the happiest country:
```{r Saudi Arabia VS Finland}
SvsF <- filter(happiness_2019, Country.or.region == "Saudi Arabia" | Overall.rank == 1)
knitr::kable(SvsF)

```
<br />

#### Piesharts to illustrate the comparisons between Saudi Arabia and Finland:
```{r Comparisons Plot, echo = FALSE}
# In plot
# Columes to get rid of
drop <- c("Country.or.region", "Overall.rank", "Score")
lab <- names(SvsF[!(names(SvsF)) %in% drop]) 

# Saudi Arabia plot
SvsF <- filter(happiness_2019, Country.or.region == "Saudi Arabia" | Overall.rank == 1)
Sa <- as.double(unlist(unname(SvsF, happiness_2019$Country.or.region == "Saudi Arabia"))) #%>%
#  select(-"Score")
Sa <- na.omit(Sa)
pie(labels = lab, Sa, main = "Saudi Arabia")

# Finland plot
Fin <- as.double(unlist(unname(SvsF[happiness_2019$Country.or.region == "Finland",])))
Fin <- na.omit(Fin)
pie(labels = lab, Fin, main = "Finland")
```

So what we need to improve in order to be in the first place are:

- `r names(SvsF)[7]` by `r SvsF[1,7] - SvsF[2,7]`
- `r names(SvsF)[4]` by `r SvsF[1,4] - SvsF[2,4]`
- `r names(SvsF)[5]` by `r SvsF[1,5] - SvsF[2,5]`
- `r names(SvsF)[6]` by `r SvsF[1,5] - SvsF[2,6]`


<br />

#### How happiness' rank in Saudi Arabia changed over the last five years:

```{r SA Happiness Rate over last 5 years, echo = FALSE}


years <- c("2015", "2016", "2017", "2018", "2019")
happinessRate <- c(sa2015$Happiness.Rank,
                   sa2016$Happiness.Rank,
                   sa2017$Happiness.Rank,
                   sa2018$Overall.rank,
                   sa2019$Overall.rank)


#Plot
#SaHappinessRate <- data.frame(years,happinessRate)
#ggplot(SaHappinessRate, aes(x = years,
#                            y = happinessRate)) +
#   geom_bar(stat="identity", fill = "#8fc0d1")

SaHappinessRate <- data.frame(years,happinessRate)
ggplot(SaHappinessRate, aes(x = years,
                            y = happinessRate)) +
   geom_point(color = "#8fc0d1")+
  theme(panel.grid = element_blank())
```
**Relation between Social Support and Ovarall.rank in 2018 and 2019:**
As it is shown Social Support does not make a huge difference in Saudis' happiness rate in the last two years.
```{r}
sociVsRank18 <- select(sa2018, c(Social.support,Overall.rank))
sociVsRank19 <- data.frame(Social.support = sa2019$Social.support,
                           Overall.rank = sa2019$Overall.rank)

sociVsRank18_19 <- rbind(sociVsRank18, sociVsRank19) 
sociVsRank18_19$year <- c("2018", "2019")
knitr::kable(sociVsRank18_19)

```


#### Saudis' happiness details over the last two years:

```{r info18_19}
Sa19 <- data.frame(Overall.rank = sa2019$Overall.rank,
                       Country.or.region = sa2019$Country.or.region,
                       Score = sa2019$Score,
                       GDP.per.capita = sa2019$GDP.per.capita,
                       Social.support = sa2019$Social.support,
                       Healthy.life.expectancy = sa2019$Healthy.life.expectancy,
                       Freedom.to.make.life.choices = sa2019$Healthy.life.expectancy,
                       Generosity = sa2019$Generosity,
                       Perceptions.of.corruption = sa2019$Perceptions.of.corruption)
                       
                       
SaHappinessDf18_19 <- rbind(sa2018, Sa19) 
SaHappinessDf18_19$year <- c("2018", "2019")
knitr::kable(SaHappinessDf18_19)
```

#### Old data..
**2015**
```{r}
knitr::kable(sa2015)
```
**2016**
```{r}
knitr::kable(sa2016)
```
**2017**
```{r}
knitr::kable(sa2017)
```


